---
title: "Download Data"
author: "Jessica Scarborough"
date: "12/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages}

library(RTCGAToolbox)
library(here)
library(tidyverse)

```

In this script, we will download all the data required for this project from GDSC and TCGA. This script will only download epithelial origin cancer samples from the TCGA. GDSC will be filtered for only epithelial cell lines in `clean_data.Rmd`.

We'll start by creating some directories to store these data. 

```{r make_dir}

dir.create(here("Data"), showWarnings = FALSE)
dir.create(here("Data", "GDSC"), showWarnings = FALSE)

```

# GDSC

## Cell Line Meta Data

First, we'll create the `get_meta` function to use `curl_download` from the `curl` package to download the cell line characterization (meta) data using file transfer protocol (FTP), storing the unaltered excel file in the `Data/GDSC` sub-directory of our project directory. We'll read this file using `read_excel` from the `readxl` package and store the data as a local variable, `data_raw_meta`. 

```{r load_meta}

get_meta <- function(cache) {
  
  if(!file.exists(paste0(cache, "/Cell_Lines_Details.xlsx"))) {
    curl::curl_download("ftp://ftp.sanger.ac.uk/pub/project/cancerrxgene/releases/current_release/Cell_Lines_Details.xlsx",
                        destfile = paste0(cache, "/Cell_Lines_Details.xlsx"))
  } else {
    message("using cached metadata")
  }
  df <- readxl::read_excel(paste0(cache, "/Cell_Lines_Details.xlsx"))
  return(df)
}

data_raw_meta <- get_meta(here("Data", "GDSC"))

```

As originally loaded, the cell line meta data set (`data_raw_meta`) has `r nrow(data_raw_meta)` rows and `r ncol(data_raw_meta)` columns. 


## Drug response

For the drug response data, we'll create the `get_dr` function, using `curl_download` from the `curl` package to download the drug response data using file transfer protocol (FTP), storing the unaltered excel file in the `Data/GDSC` sub-directory of our project directory. We'll read this file using `read_excel` from the `readxl` package and store the data as a local variable, `gdsc_dr`. 

```{r load_dr}
options(timeout = 600) # increased from default of 60 s

get_dr <- function(cache) {
  
  # check if file has already been downloaded at the provided cache
  if(!file.exists(paste0(cache, "/gdsc_dr.xlsx"))) {
    
    # download file, store in `Data` dir
    curl::curl_download(
      "ftp://ftp.sanger.ac.uk/pub/project/cancerrxgene/releases/current_release/GDSC2_fitted_dose_response_25Feb20.xlsx",
      destfile = paste0(cache, "/gdsc_dr.xlsx"))
    
  } else {
    message("using cached gdsc data")
  }
  
  #read into R
  df <- readxl::read_excel(paste0(cache, "/gdsc_dr.xlsx"))
  return(df)
  
}

gdsc_dr <- get_dr(here("Data", "GDSC"))


```

As originally loaded, the drug response data set (`gdsc_dr`) has `r nrow(gdsc_dr)` rows and `r ncol(gdsc_dr)` columns.


## Amp/Deletion data

For the deletion (and amplification data), we'll create the `get_del` and `get_del_map` functions, using `curl_download` from the `curl` package to download the deletion event and mapping data using file transfer protocol (FTP), storing the unaltered excel file in the `Data/GDSC` sub-directory of our project directory. We'll read this file using `read_excel` from the `readxl` package and store the data as a local variable, `gdsc_del` and `gdsc_del_map`. 

The mapping is necessary because the deletion events are denoted by "region identifiers" such as cnaLUAD1 or cnaGBM81. These identifiers are then mapped to chromosomal locations in a separate file. We will extract only the deletions of interest in the `clean_data.Rmd` file. 

### Amp/Deletion events

```{r load_del}
options(timeout = 600) # increased from default of 60 s

get_del <- function(cache) {
  
  # check if file has already been downloaded at the provided cache
  if(!file.exists(paste0(cache, "/gdsc_del.xlsx"))) {
    
    # download file, store in `Data` dir
    curl::curl_download(
      "https://www.cancerrxgene.org/gdsc1000/GDSC1000_WebResources//Data/suppData/TableS2G.xlsx",
      destfile = paste0(cache, "/gdsc_del.xlsx"))
    
  } else {
    message("using cached gdsc data")
  }
  
  #read into R
  df <- readxl::read_excel(paste0(cache, "/gdsc_del.xlsx"))
  return(df)
  
}


gdsc_del <- get_del(here("Data", "GDSC"))


```

### Amp/Deletion mapping

```{r load_del_map}
options(timeout = 600) # increased from default of 60 s

get_del_map <- function(cache) {
  
  # check if file has already been downloaded at the provided cache
  if(!file.exists(paste0(cache, "/gdsc_del_chr_map.xlsx"))) {
    
    # download file, store in `Data` dir
    curl::curl_download(
      "https://www.cancerrxgene.org/gdsc1000/GDSC1000_WebResources//Data/suppData/TableS2D.xlsx",
      destfile = paste0(cache, "/gdsc_del_chr_map.xlsx"))
    
  } else {
    message("using cached gdsc data")
  }
  
  #read into R
  df <- readxl::read_excel(paste0(cache, "/gdsc_del_chr_map.xlsx"))
  return(df)
  
}


gdsc_del_map <- get_del_map(here("Data", "GDSC"))

```
